A Framework for Classification of Social Machines

### 2013-02-15 latex/sig-alternate.tex
Methodology added

* research community: shared terminology, set of key features and their possible instantiations based on the analysis of a relevant set of 50 (?) social machines (both popular systems as well as long tail systems); facilitates the study of design and evolution patterns; supports research on how communities emerge and develop and organizational studies (mechanism design, incentives); different research approaches can be better compared and aligned.

* developer community: provides a theoretical grounding for the realization of new systems, informs system design and operation (what features should be implemented, what features are more important than others); enables predictions on user behavior and community development.

@@

How was the classification framework defined

* interviews with XXX researchers on key descriptive features of systems they are familiar with

* consolidation of features listed for a resulting set of 50 (?) social machines

* grid analysis to identify groups of similar systems

* evaluation with respect to correctness, completeness, clarity, extendability in user survey, additional data set of XXX systems

@@

First explain main principles: Distinction between application system and enabling technology; contributions, actions, activities, purpose, etc

\subsection{The Hierarchy/Polyarchical relationship of social machines}

@@

Looking at the set of social machines in a polyarchy leads to a broad/specific relationship emerging that lets us talk about behaviours at various levels of granularity.  We propose looking at nested machines with ``The Web'' as one of several potential roots, with the next level down consisting of sub-platforms (such as Facebook, Twitter, MediaWiki, Ushahid platforms) that spawn more specific social machines. A resulting polyarchy is shown in Figure~\ref{polyarchy}, which classifies levels as Infrastructure, Frameworks, Services and Causes/Groups.

@@

\section{Usage examples (0.5 page)}
Wikipedia

>>>>

* research community: shared terminology, set of key features and their possible instantiations based on the analysis of a relevant set of 50 (?) social machines (both popular systems as well as long tail systems); facilitates the study of design and evolution patterns; supports research on how communities emerge and develop and organisational studies (mechanism design, incentives); different research approaches can be better compared and aligned.

* developer community: provides a theoretical grounding for the realisation of new systems, informs system design and operation (what features should be implemented, what features are more important than others); enables predictions on user behaviour and community development.

@@

%How was the classification framework defined
%
%* interviews with XXX researchers on key descriptive features of systems they are familiar with
%
%* consolidation of features listed for a resulting set of 50 (?) social machines
%
%* grid analysis to identify groups of similar systems
%
%* evaluation with respect to correctness, completeness, clarity, extendability in user survey, additional data set of XXX systems
%

In order to identify and classify social machines, we first embarked on a principled
knowledge elicitation process~\cite{knowledgeelicitation}, to determine the {\it elements}
(i.e., the social machines) and the {\it constructs} (i.e., the classifying factors) that
we can use to classify social machines. We used the {\it repertory grid} elicitation
technique~\cite{kelly} in order to derive an initial set of elements and constructs. We
asked ten Computer Science researchers familiar with social machines to create their own
repertory grids, and populate the elements from their own knowledge, and to create the
constructs using the standard repertory grid elicitation technique. This exercise led
to 10 grids, the union of which comprised a total of 56 unique elements (social machines)
and 117 different constructs.

To make the set of social machines easier to refer to and more approachable, we performed
an initial grid analysis of the machines, clustering them into related user-centered
themes, of: Social Networks, (Micro)Blogging, News Aggregators, Image Boards, Crowd Science,
Answer Gardens, Community Watch, Health and Wellbeing Support, Action and Investigation,
Opinion Sharing, Video Sharing, Photo Sharing, Code Sharing, Art Sharing, Crowdsourcing
Platforms, Mash-up Systems, Crowdsourcing Toolkits and Platforms.

While determining the intersection of elements was straightforward, the consolidation of the
constructs required a more thorough process. To consolidate the constructs we embarked on a manual process of grouping the constructs
into rough clusters, based around the areas they cover. For example, we used construct
clusters including: Users, Motivation, Popularity, Technology, and Purpose. We then examined
each construct to determine which were equivalent, and whether we could re-word or subsume
existing constructs to cover the same aspects as a construct, with the aim to remove any
redundancy, to cut the list of constructs down to a manageable number, but to ensure that
all of the constructs that were elicited were somehow represented in the final set. This
process involved four of the authors discussing the choices, and finally agreeing on a
consolidated set of 31 constructs.

Even with this reduced set of constructs, if a single person were to classify all 56
elements, this would still require 1736 individual decisions, and hence is unfeasible to
expect full coverage of the elements using all of these constructs, despite how representative
they are.

[something here about doing 2 grids each and verifying the results?]

To extract greater usefulness, and the ability to use our constructs to classify social
machines we re-evaluated the constructs and the semantics that they represent into richer
constructs that are not constrained by the 1--5 rating of repertory grid constructs.

@@

%First explain main principles: Distinction between application system and enabling technology; contributions, actions, activities, purpose, etc

To create a classification framework, it is necessary to first determine the aspects of
classification. We will first discuss the relationships between social
machines, enabling technologies and platforms; we move on to discuss the multi-faceted
nature of motivation within social machines; next we evaluate the artefacts of social
machines: contributions, actions, activities and purpose; finally we discuss the metrics of
participation, and the difference between designed and emergent properties of social
machines.

\subsection{The Hierarchical/Polyarchical relationship of social machines}

@@

Looking at the set of social machines in a polyarchy leads to a broad/specific relationship emerging that lets us talk about behaviours at various levels of granularity.  We propose looking at nested machines with ``The Web'' as one of several potential roots, with the next level down consisting of sub-platforms (such as Facebook, Twitter, MediaWiki, Ushahidi platforms) that spawn more specific social machines. A resulting polyarchy is shown in Figure~\ref{polyarchy}, which classifies levels as Infrastructure, Frameworks, Services and Causes/Groups.

@@

\section{Usage examples (0.5 page)}

Wikipedia
